PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-05 paper/sec-doublecounting.tex
small edits in sec-doublecounting

\begin{figure}
    \centering
    \begin{overpic}[width=0.6\textwidth, trim=0.cm 11cm 0.cm 10cm, clip=True]{figures/dy-pi}
        \put (5, 33) {\large $\mathrm{q}$}
        \put (20, 37) {\large $\mathrm{q}$}
        \put (38, 5) {\large $\bar{\ell}$}
        \put (38, 31) {\large $\ell$}
        \put (11, 6) {\large $\gamma$}
        %
        \put (65, 9) {\large $\mathrm{q}$}
        \put (73, 30) {\large $\mathrm{q}$}
        \put (96, 25) {\large $\ell$}
        \put (96, 8) {\large $\bar{\ell}$}
        \put (55, 26) {\large $\gamma$}
    \end{overpic}
    \caption{\label{fig:dy-pi}
    Photon-induced (left) and quark-induced (right) contributions to the Drell-Yan process. In black, the LO process is shown.
    In red, the initial-state splitting leading to the real-emission $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ is highlighted. Such a
    real emission enters in the NLO EW corrections.}
\end{figure}
An example is the subtraction of (irreducible) background processes which must not be considered as such. A very blatant case
is neutral-current Drell--Yan, where the signal process is the production of an opposite-sign lepton pair, which starts
at $\mathcal O(\alpha^2)$. Because this process is sometimes thought
as a quark-initiated $s$-channel mechanism ($\mathrm{q} \bar{\mathrm{q}} \to \gamma^*/\mathrm{Z} \to \ell \bar{\ell}$), in many analyses the photon-induced component,
$\gamma \gamma \to \ell \bar{\ell}$ in the $t$ channel, is considered to be a different process, and therefore as a background and subtracted.
The subtraction from the measured data is done by calculating the theoretical predictions of the double-photon initiated contribution, possibly including (unphysical) higher-order
corrections. For example, in refs.~\cite{Aaboud:2017ffb,Aad:2016zzw} (a similar statement appears also in an older analysis~\cite{Aad:2013iua}), one reads:
\begin{quote}
The photon-induced process, $\gamma\gamma \to \ell \bar{\ell}$, is simulated at LO using Pythia 8
and the MRST2004qed PDF set~\cite{Martin:2004dh}. The expected yield for this process also accounts for 
NLO QED/EW corrections from references~\cite{Bardin:2012jk,Bondarenko:2013nu}, which decrease the yield by approximately \SI{30}{\percent}.
\end{quote}
Such a distinction, which is unphysical and incorrect in quantum mechanics, may be somehow justified at LO. Beyond this order, it is simply wrong.
Indeed, at $\mathcal O(\alpha^3)$, the reaction $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ becomes possible, which
includes both kind of topologies discussed above (and needs both in order to yield an IR-finite result), see Fig.~\ref{fig:dy-pi}. While it is common sense that the QCD counterpart of this subtraction should never 
be performed --- no one would
consider to \enquote{subtract} the gluon-initiated contribution to top-pair production in top analyses --- seemingly it is not so
when EW corrections are considered.

@@

It is interesting
to note that one can tune the QED parton shower to specific decays to mimic NLO EW effects, so that a prediction only accurate at NLO
QCD displays a remarkable agreement with another at NLO QCD+EW when the photonic shower is included (see for example the behaviour of predictions showered with
\textsc{Photos}~\cite{Barberio:1990ms,Barberio:1993qi,Golonka:2005pn} in ref.~\cite{CarloniCalame:2016ouw}). However, this kind of agreement
always comes \emph{a posteriori}, and cannot be ensured in general.
Another problem with this approach is that it introduces a dependence on the program that is used for the shower inversion, or even on a specific version of the program, for examples if bugs have been found.
This is especially problematic if these datasets are the only ones published, since undoing the exact subtraction can be very difficult or practically impossible as time passes (TODO insert comment about LEP ISR and DIS EW effects removal).

>>>>

\begin{figure}
    \centering
    \begin{overpic}[width=0.6\textwidth, trim=0.cm 11cm 0.cm 10cm, clip=True]{figures/dy-pi}
        \put (5, 33) {\large $\mathrm{q}$}
        \put (20, 37) {\large $\mathrm{q}$}
        \put (38, 5) {\large $\bar{\ell}$}
        \put (38, 31) {\large $\ell$}
        \put (11, 6) {\large $\gamma$}
        %
        \put (65, 9) {\large $\mathrm{q}$}
        \put (73, 30) {\large $\mathrm{q}$}
        \put (96, 25) {\large $\ell$}
        \put (96, 8) {\large $\bar{\ell}$}
        \put (55, 26) {\large $\gamma$}
    \end{overpic}
    \caption{\label{fig:dy-pi}
    Photon-induced (left) and quark-induced (right) contributions to the Drell-Yan process. In black, the LO process is shown.
    In red, the initial-state splitting leading to the real-emission $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ is highlighted. Such a
    real emission enters in the NLO EW corrections.}
\end{figure}
An example is the subtraction of (irreducible) background processes which must not be considered as such. A very blatant case
is neutral-current Drell--Yan, where the signal process is the production of an opposite-sign lepton pair, which starts
at $\mathcal O(\alpha^2)$. Because this process is usually thought
as a quark-initiated $s$-channel mechanism ($\mathrm{q} \bar{\mathrm{q}} \to \gamma^*/\mathrm{Z} \to \ell \bar{\ell}$), in many analyses the photon-induced component,
$\gamma \gamma \to \ell \bar{\ell}$ in the $t$ channel, is considered to be a different process, and therefore as a background and subtracted.
The subtraction from the measured data is done by calculating the theoretical predictions of the double-photon initiated contribution, possibly including (ill-defined) higher-order
corrections. For example, in refs.~\cite{Aaboud:2017ffb,Aad:2016zzw} (a similar statement appears also in an older analysis~\cite{Aad:2013iua}), one reads:
\begin{quote}
The photon-induced process, $\gamma\gamma \to \ell \bar{\ell}$, is simulated at LO using Pythia 8
and the MRST2004qed PDF set~\cite{Martin:2004dh}. The expected yield for this process also accounts for 
NLO QED/EW corrections from references~\cite{Bardin:2012jk,Bondarenko:2013nu}, which decrease the yield by approximately \SI{30}{\percent}.
\end{quote}
Such a distinction, which is unphysical and incorrect in quantum mechanics, may be somehow justified at LO. Beyond this order, it is simply wrong.
Indeed, at $\mathcal O(\alpha^3)$, the reaction $\mathrm{q} \gamma \to \ell \bar{\ell} \mathrm{q}$ becomes possible, which
includes both kind of topologies discussed above, and needs both in order to yield an IR-finite result, see Fig.~\ref{fig:dy-pi} (as a consequence, one cannot speak of EW corrections to $\gamma \gamma \to \ell \bar{\ell}$). While it is common sense that the QCD counterpart of this subtraction should never 
be performed --- no one would
consider to \enquote{subtract} the gluon-initiated contribution to top-pair production in top analyses --- seemingly it is not so
when EW corrections are considered.

@@

It is interesting
to note that one can tune the QED parton shower to mimic NLO EW effects for specific processes and observables, so that a prediction only accurate at NLO
QCD displays a remarkable agreement with another at NLO QCD+EW when the photonic shower is included (see for example the behaviour of predictions showered with
\textsc{Photos}~\cite{Barberio:1990ms,Barberio:1993qi,Golonka:2005pn} in ref.~\cite{CarloniCalame:2016ouw}). However, this kind of agreement
always comes \emph{a posteriori}, and cannot be ensured in general.
Furthermore, the deconvolution of QED effects in data introduces a dependence on the program (and possibly on 
the specific version) employed for the shower inversion. 
This fact is especially problematic if deconvolved datasets are the only ones which are published, since undoing the exact deconvolution can be very difficult, or practically impossible, as time passes. (TODO insert comment about LEP ISR and DIS EW effects removal).
